---
title: "A tour of GeneCTLR"
author: "Kara Han"
date: "`r format(Sys.time(), '%d %b %Y')`"
output: 
  rmarkdown::html_vignette:
    toc: true
    number_sections: false
vignette: >
  %\VignetteIndexEntry{A tour of GeneCTLR}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
library(knitr)
opts_chunk$set(fig.align = "center", 
               out.width = "90%",
               fig.width = 6, fig.height = 5.5,
               dev.args=list(pointsize=10),
               par = TRUE, # needed for setting hook 
               collapse = TRUE, # collapse input & ouput code in chunks
               warning = FALSE)
knit_hooks$set(par = function(before, options, envir)
  { if(before && options$fig.show != "none") 
       par(family = "sans", mar=c(4.1,4.1,1.1,1.1), mgp=c(3,1,0), tcl=-0.5)
})
set.seed(1029) # for exact reproducibility
```

## Get started
`GeneCTLR` is an R package designed to provide a workflow for logistic
regression and gene classifications. In this vignette, we present an example introductory workflow for predicting autism causal genes with a logistic regression model. 
**This document gives a tour of GeneCTLR (version 0.1.0)**. 
It was written in R Markdown, using the [knitr](https://cran.r-project.org/package=knitr) package for production. See `help(package = "GeneCTLR")` for further details.

To download **GeneCTLR**, use the following commands:
```{r}
require("devtools")
devtools::install_github("karahyx/GeneCTLR")
library("GeneCTLR")
```

To list all sample functions available in the package:
```{r}
ls("package:GeneCTLR")
```

To list all sample datasets available in the package:
```{r}
data(package = "GeneCTLR")
```

## Load in the data
Here, we analyze a data set of 1072 RNA-binding proteins (RBPs) annotated for a 
range of features, including neuron enrichment, brain enrichment, neuron 
developmental enrichment, pLI scores, whether the RBP contains a canonical 
RNA-binding protein (RBD), and whether the RBP is associated with autism based 
on data obtained from the Simons Foundation Autism Research Initiative (SFARI). 
The **autism_genes** variable is the dependent variable in the model.

The rbps data set is a data frame that contains 1079 observations of 12 
variables (Human Gene, hasCanonicalRBDs, Neuron Enrichment (Mouse), Neuron 
Enrichment (Human), Brain Enrichment, day3/day1_diff, day7/day1_diff, pLI, 
medium_pLI, high_pLI, extreme_pLI, and autism_genes). 
To load in the rbps data set:
```{r}
data(rbps)
head(rbps)
```

## Workflow
### Checking for missing values
To prepare the rbps data set for our analysis, we need to first check whether there exist missing or corrupted data and how many unique values there are for each variable.
```{r, echo = TRUE}
missingMap <- GeneCTLR::missingValues(rbps)
```

### Taking care of the missing values (data imputation)
Missing values in the data set can cause problems in our future analysis, so we need to account for the `NAs` by replacing them with the mean or the median of each variable. Here, we will be using the mean for the rbps data set.
```{r, echo = TRUE}
imputedRBPs <- GeneCTLR::impute(rbps, "mean")
head(imputedRBPs)
```

### Model fitting
After data cleaning, we are ready to fit the model. The `trainCV()` function 
builds the logistic regression model with K-fold cross validation, and returns 
a list object with `models`, `predictions`, and `testSets`. 
``` {r, echo = TRUE}
# Since the trainCV function only works with one dependent variable of type
# factor and independent variables of type numeric, we will use preProcessData()
# to pre-process the data set for trainCV().
newDat <- GeneCTLR::preProcessData(data = imputedRBPs,
                                   dependentVarIndex = 12,
                                   deleteColumns = c(1, 8))

# Model fitting
# The autism_genes variable is the 10th column in newDat, so dependentVarIndex = 10
# Using default K = 5
results <- GeneCTLR::trainCV(data = newDat, dependentVarIndex = 10)

results$models
```

### Plotting ROC and Precision-Recall Curves
The receiver operating characteristic curve, or ROC curve, is a plot used to demonstrate the diagnostic ability of a logistic regression model by showing the trade-off between the true positive rate and false positive rate using different probability thresholds. The precision-recall curve illustrates the trade-off between precision and recall for different probability thresholds. To plot the ROC and the precision-recall curves in R, use `plotROC()` and `plotPR()`.
```{r, echo = TRUE}
rocCurve <- GeneCTLR::plotROC(results = results,
                              dependentVarIndex = 10,
                              title = "ROC Curves for Autism Gene Prediction")
rocCurve

prCurve <- GeneCTLR::plotPR(results = results,
                            dependentVarIndex = 10,
                            title = "Precision-Recall Curves for Autism Gene Prediction")
prCurve
```


## Package References
[Han, K. (2021) GeneCTLR (Gene Classification Tool with Logistic Regression):
an R package for predicting disease gene classifications with logistic regression. An Assessment For BCB410H. Unpublished.](https://github.com/karahyx/GeneCTLR)
  
<br>

## Other References
Alice, M. (2020, July 5). *How to perform a logistic regression in R: R-bloggers.* R. https://www.r-bloggers.com/2015/09/how-to-perform-a-logistic-regression-in-r/.

Boyd, K., Eng, K. H., &amp; Page, C. D. (2013). Area under the precision-recall curve: Point estimates and confidence intervals.
*Advanced Information Systems Engineering*, 451–466. https://doi.org/10.1007/978-3-642-40994-3_29.

Bradley, A. P. (1997). The use of the area under the ROC curve in the evaluation of machine learning algorithms. *Pattern Recognition*, 30(7), 1145–1159. https://doi.org/10.1016/s0031-3203(96)00142-2.

Fu, G. H., Yi, L. Z., & Pan, J. (2018). Tuning model parameters in class‐imbalanced learning with precision‐recall curve. *Biometrical Journal*, 61(3), 652–664. https://doi.org/10.1002/bimj.201800148

Little, JA. R. & Rubin B. D. (1987). *Statistical analysis of missing data*. John Wiley & Sons.

Honaker, J., King, G., & Blackwell, M. (2011). Amelia II: A Program for Missing Data. *Journal of Statistical Software*, 45(7), 1–47. https://doi.org/10.18637/jss.v045.i07.

Google. (n.d.). *Classification: ROC curve and AUC; machine learning crash course; google developers*. Google. https://developers.google.com/machine-learning/crash-course/classification/.

Grolemund, G. (2015). Learn Shiny - Video Tutorials. https://shiny.rstudio.com/tutorial/.

Selva, P. (n.d.). *Logistic Regression with R.* r-statistics.co. http://r-statistics.co/Logistic-Regression-With-R.html.

Williams, C. K. (2021). The effect of class imbalance on precision-recall curves. *Neural Computation*, 33(4), 853–857. https://doi.org/10.1162/neco_a_01362.

Zach. (2021, September 9). *How to calculate AUC (area under curve) in R.* Statology. https://www.statology.org/auc-in-r/.

----

```{r}
sessionInfo()
```
